\subsection[cast]{Type casting}

\begin{frame}
  \frametitlecpp[98]{Type casting}
  \begin{block}{5 types of casts in \cpp\ - These 2 should be used}
    \begin{itemize}
    \item \mintinline{c++}{static_cast<Target>(arg)}
      \begin{itemize}
      \item converts type if the static types allow it
      \item including using implicit conversion
      \item essentially compile time
      \end{itemize}
    \item \mintinline{c++}{dynamic_cast<Target>(arg)}
      \begin{itemize}
      \item checks if object at address of ``\mintinline{c++}{arg}'' is convertible to \mintinline{c++}{Target}
      \item throws \mintinline{c++}{std::bad_cast} or returns  \mintinline{c++}{nullptr} if it's not
      \item essentially run time
      \end{itemize}
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}[fragile]
  \frametitlecpp[98]{Type casting example}
  \small
  \begin{exampleblock}{}
    \begin{cppcode*}{gobble=2}
      struct A{ virtual ~A()=default } a;
      struct B : A {} b;

      A& c = static_cast<A&>(b); // OK. b is also an A
      B& d = static_cast<B&>(a); // UB: a is not a B
      B& e = static_cast<B&>(c); // OK. c is a B

      B& f = dynamic_cast<B&>(c); // OK, c is a B
      B& g = dynamic_cast<B&>(a); // Exception: not a B

      B* d = dynamic_cast<B*>(&a); // nullptr. a not a B.
      if (d != nullptr) {
        // Will not reach this
      }

      if (auto bPtr = dynamic_cast<B*>(&c)) {
        // OK, we will get here
      }
      \end{cppcode*}
  \end{exampleblock}
\end{frame}

\begin{frame}[fragile]
  \frametitlecpp[98]{Type casting}
  \begin{block}{5 types of casts in \cpp - These 3 should be avoided}
    \begin{itemize}
    \item \mintinline{c++}{const_cast<Target>(arg)}
      \begin{itemize}
      \item removes constness from a type
      \item if you think you need this, rather improve your design!
      \end{itemize}
    \item \mintinline{c++}{reinterpret_cast<Target>(arg)}
      \begin{itemize}
      \item changes type irrespective of what `arg` is
      \item almost never a good idea!
      \end{itemize}
    \item C-style: \mintinline{c++}{(Target)arg}
      \begin{itemize}
      \item Force-changes type in C-style. No checks. Don't use.
      \end{itemize}
    \end{itemize}
  \end{block}
  \begin{alertblock}{Casts to avoid}
    \scriptsize
    \begin{cppcode*}{gobble=2}
      void func(A const & a) {
        A& ra = a;                 // Error: not const
        A& ra = const_cast<A&>(a); // Compiles. Bad design!
        // Evil! Don't do this:
        B* b = reinterpret_cast<B*>(&a);
        B* b = (B*)&a;
      }
    \end{cppcode*}
  \end{alertblock}
\end{frame}
